---
title: "Final Homework 5"
author: "Benise Limon"
format: 
  html: 
    toc: true 
    toc-location: left
    code-fold: true 
    theme: yeti 
editor: visual
execute: 
  message: false
  warning: false
---

## Preparation

```{r load-packages}
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
library(dplyr)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```

## Introduction

Write a 5-10 sentence introduction to the analysis you conduct. Some useful information\
could include (but is not limited to):\
a. What are Sarracenia?

Sarracenia are a pitcher plant that is a genus of carnivorous plants native to the Eastern United States, Texas, and the Great Lakes. It is typically found in nutrient-poor environments and has developed the carnivorous capabilities to supplement its nutrient intake (Hotti et al. 2017).

\
b. Why might people care about them?

Humans may care about sarracenia because it provides medical services to humans because its extractions can be used to treat type 2 diabetes and tuberculosis-like symptoms (Podolak et al. 2022)

\
c. Why do you think it could be useful and/or important to predict individual biomass from morphological, physiological, or taxonomic characteristics?

Since sarracenia is very beneficial to humans, growing it in large amounts is ideal. It can be useful to know what morphological, physiological, or taxonomic characteristics can predict the largest amounts of individual biomass.

\
d. What questions are you addressing?

In this analysis, I hope to address what morphological, physiological, or taxonomic characteristics are most influential to predict individual biomass. Various characteristics include chlorophyll, mass, specific leaf area, phyllodes, and pitchers and number of phyllodes.

\
e. What hypotheses are you testing?

\
f. Include 3-5 in-text citations of peer-reviewed work.

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5319649/

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5319649/

## Methods

Write a methods section that includes\
a. Sampling methods (these should be summarized from the metadata and cited)\
b. Description of data organization and processing (1-3 sentences)

```{r reading-data}
plant <- read.csv(here("data/knb-lter-hfr/hf109-01-sarracenia.csv")) %>%
  #making column names cleaner 
  clean_names() %>% 
  #selecting the columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```

\
c. Visualization of missing observations with an accompanying caption and discussion

```{r missing-data-vis, fig.cap = "Figure 1: Visualization of missing data observations. Chlorophyll, Mass, Specific Leaf Area, Phyllodes, and Pitchers and Phyllodes data is missing."}
#visualizing missing observations 
gg_miss_var(plant)+
  #adding a caption 
#  labs( caption = "Visualization of missing data observations. Chlorophyll, Mass, Specific Leaf Area, Phyllodes, and Pitchers and Phyllodes data is missing.") +
  theme(plot.caption = element_text(size=10, hjust = 0.8, vjust = 2))
```

```{r subset-data}
#create a subset of the plant dataframe that does not include missing values 
plant_subset <- plant %>% 
  drop_na(sla, chlorophyll, amass, num_lvs, num_phylls)
```

\
Discussion: 1-3 sentences

\
d. Visualization of Pearson's correlation between variables with an accompanying\
caption and discussion.

```{r pearson, fig.cap = "Figure 2: A correlation plot determining the relationship between variables in the dataset. The higher the number, the higher correlation."}
# calculate Pearson's r for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")
  
# creating a correlation plot
corrplot(plant_cor,
         # change the shape of what's in the cells
         method = "ellipse",
         addCoef.col = "black"
         )
```

Discussion: 1-3 sentences

\
e. Visualization of relationships between variables (not correlation) with an\
accompanying caption and discussion (caption: 1-3 sentences, discussion: 1-3\
sentences)

\
f. Creating the null and full model with accompanying discussion (1-3 sentences)\
g. Visual and statistical assumption checks for the full model with accompanying\
discussion (1-3 sentences)\
h. Explanation of any transformation and justification (1-3 sentences)\
i. Model construction with visual and statistical assumption checks for three\
additional models with justification (i.e. why did you choose those predictor\
variables?) (justification: 1-3 sentences per model)\
j. Variance inflation factor check for the full model with accompanying discussion\
(discussion: 1-3 sentences)\
k. Model comparison using the null model, full model, and additional models from part 7g

## Results

Write a results section that includes\
a. A description of the best model and how you chose it, with a summary of model\
results (parenthetical and/or table, 1-3 sentences)\
b. Interpretation of your chosen model (1-3 sentences)\
c. Visualization of model predictions for biomass as a function of the predictor\
variable of your choice with an accompanying caption (caption: 1-3 sentences)\
d. Discussion of results: what does this model mean biologically? (1-3 sentences)

## Bibliography

Ellison, Aaron M., Hannah L. Buckley, Thomas E. Miller, and Nicholas J. Gotelli. 2004. "Morphological Variation in *Sarracenia Purpurea* (Sarraceniaceae): Geographic, Environmental, and Taxonomic Correlates." *American Journal of Botany* 91 (11): 1930--35. <https://doi.org/10.3732/ajb.91.11.1930>.
