---
title: "Final Homework 5"
author: "Benise Limon"
format: 
  html: 
    toc: true 
    toc-location: left
    code-fold: true 
    theme: yeti 
editor: visual
execute: 
  message: false
  warning: false
---

## Preparation

```{r load-packages}
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
library(dplyr)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```

## Introduction 

Write a 5-10 sentence introduction to the analysis you conduct. Some useful information\
could include (but is not limited to):\
a. What are Sarracenia?\
b. Why might people care about them?\
c. Why do you think it could be useful and/or important to predict individual biomass from morphological, physiological, or taxonomic characteristics?\
d. What questions are you addressing?\
e. What hypotheses are you testing?\
f. Include 3-5 in-text citations of peer-reviewed work.

## Methods

Write a methods section that includes\
a. Sampling methods (these should be summarized from the metadata and cited)\
b. Description of data organization and processing (1-3 sentences)

```{r reading-data}
plant <- read.csv(here("data/knb-lter-hfr/hf109-01-sarracenia.csv")) %>%
  #making column names cleaner 
  clean_names() %>% 
  #selecting the columns of interest 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```

\
c. Visualization of missing observations with an accompanying caption and discussion

```{r missing-data-vis, fig.cap = "Visualization of missing data observations. Chlorophyll, Mass, Specific Leaf Area, Phyllodes, and Pitchers and Phyllodes data is missing."}
#visualizing missing observations 
gg_miss_var(plant)+
  #adding a caption 
#  labs( caption = "Visualization of missing data observations. Chlorophyll, Mass, Specific Leaf Area, Phyllodes, and Pitchers and Phyllodes data is missing.") +
  theme(plot.caption = element_text(size=10, hjust = 0.8, vjust = 2))
```

```{r subset-data}
#create a subset of the plant dataframe that does not include missing values 
plant_subset <- plant %>% 
  drop_na(sla, chlorophyll, amass, num_lvs, num_phylls)
```

\
(caption: 1-3 sentences, discussion: 1-3 sentences)\
d. Visualization of Pearson\'s correlation between variables with an accompanying\
caption and discussion (caption: 1-3 sentences, discussion: 1-3 sentences)

\
e. Visualization of relationships between variables (not correlation) with an\
accompanying caption and discussion (caption: 1-3 sentences, discussion: 1-3\
sentences)\
f. Creating the null and full model with accompanying discussion (1-3 sentences)\
g. Visual and statistical assumption checks for the full model with accompanying\
discussion (1-3 sentences)\
h. Explanation of any transformation and justification (1-3 sentences)\
i. Model construction with visual and statistical assumption checks for three\
additional models with justification (i.e. why did you choose those predictor\
variables?) (justification: 1-3 sentences per model)\
j. Variance inflation factor check for the full model with accompanying discussion\
(discussion: 1-3 sentences)\
k. Model comparison using the null model, full model, and additional models from part 7g

## Results 

Write a results section that includes\
a. A description of the best model and how you chose it, with a summary of model\
results (parenthetical and/or table, 1-3 sentences)\
b. Interpretation of your chosen model (1-3 sentences)\
c. Visualization of model predictions for biomass as a function of the predictor\
variable of your choice with an accompanying caption (caption: 1-3 sentences)\
d. Discussion of results: what does this model mean biologically? (1-3 sentences)

## Bibliography

Ellison, Aaron M., Hannah L. Buckley, Thomas E. Miller, and Nicholas J. Gotelli. 2004. \"Morphological Variation in *Sarracenia Purpurea* (Sarraceniaceae): Geographic, Environmental, and Taxonomic Correlates.\" *American Journal of Botany* 91 (11): 1930--35. <https://doi.org/10.3732/ajb.91.11.1930>.
